Interactively Visualizing Multivariate Market Segmentation Using pytourr

### 2024-09-26 manuscript.tex
changed figure descriptions and added color indicators to the main text

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{silhouette_comparison.png}
    \caption{A: silhouette plot of the k-means solution of the full dataset. B: silhouette plot of the k-means solution with the reduced dataset.}
    \label{fig:silhouette_comparison}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_activ_cluster_highlights.png}
    \caption{Cluster highlights: \textbf{A} - Cluster 1 (blue), \textbf{B} - Cluster 2 (orange), \textbf{C} - Cluster 3 (green), \textbf{D} - Cluster 4 (red), \textbf{E} - Cluster 5 (violet), \textbf{F} - Cluster 6 (brown).}
    \label{fig:winter_activ_cluster_highlights}
\end{figure}

This process allows us to visually assess the separation and similarities between the clusters, providing insight into the structure of the dataset. By highlighting each cluster individually, we can evaluate their distinctiveness in different projections. The most influential features shown in Figure \ref{fig:winter_activ_cluster_highlights} are pool/sauna, alpine skiing, museums, and going to the spa. The projection roughly separates clusters 1, 2, and 3 from each other and the other three clusters, which appear to be quite similar.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_init.png}
    \caption{Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot.}
    \label{fig:winter_cl7_init}
\end{figure}

@@

This setup produces the display shown in Figure \ref{fig:winter_cl7_init}. It can also be seen that both clusters 1 and 3 contain tourists that didn't go alpine skiing and the main difference between them is that tourists in cluster 3 enjoyed going to the spa and health facilities as well as going to the pool, while the ones in cluster 1 didn't. Other than that, the clusters were mostly similar. Now we might be interested in the subset of tourists that enjoy going to museums. Therefore, we can adjust the projection axes so that these axes are elongated and point into different directions. We can see that there is indeed overlap between clusters 1 and 3, as shown in Figure \ref{fig:winter_cl7_pre}. As a next step we can reassign the overlapping section to a new cluster - cluster 7. By selecting the checkbox for subset 7 and manually selecting the region of overlap, we can form a new cluster, which is visualized in Figure \ref{fig:winter_cl7_post}.

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_pre.png}
    \caption{Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot.}
    \label{fig:winter_cl7_pre}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_post.png}
    \caption{Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot.}
    \label{fig:winter_cl7_post}
\end{figure}

@@

At first, hierarchical clustering using Ward's method \citep{ward2} and the Jaccard index was applied to the features. The resulting dendrogram is shown in Figure \ref{fig:aus_feature _clustering}. Based on this clustering, \( k = 15 \) clusters were identified, and generally, only one representative feature from each cluster was selected for further analysis. Clusters containing unpopular activities, such as "Adventure", which only had 42 participants, were discarded. The cluster containing popular features like "Beach," "Swimming," "ScenicWalks," "Markets," "Sightseeing," "Friends," "Pubs," "BBQ," "Shopping," "Eating," "EatingHigh," "Movies," and "Relaxing" was treated differently. Multiple features from this cluster were retained to preserve as much information as possible. After feature selection, the observations were reclustered using k-means with \( k = 15 \).

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{aus_feature_clustering.png}
    \caption{Dendogram of the features of the Australian Vacation Activities dataset using Ward's method with the Jaccard index.}
    \label{fig:aus_feature_clustering}
\end{figure}

@@

Since we are particularly interested in tourists who did not spend time with friends, we can extend the projection axis "Friends" outward to separate the data based on that feature. Datapoints in the opposite direction of the "Friends" axes are the solo travellers we are interested in, which can be seen in Figure \ref{fig:aus_preselection} on the left side of the top left plot. Subsequently, we can pull all other projection axes in one direction to separate data points based on their overall activity level. Tourists that fall in the direction of the axes generally engaged in more activities compared to those in the opposite direction of the projection axes. This separation is evident, as observations in subset 6 are located opposite the axes, and the heatmap shows that they did not engage in many activities. Similarly we can also see that subset 2, which contains quite active tourists, is shifted towards the direction of the axes.

Using this logic, we can segment the data points on the left into three segments: active tourists (more upward), medium active tourists (center), and largely inactive tourists (bottom). The result of this segmentation can be seen in Figure \ref{fig:aus_selection}. Analyzing the heatmap, we can identify several interesting patterns.

By comparing subset 2, which contains active tourists who spent time with their friends, with subset 7, the active tourists who traveled alone, we notice that subset 7 showed less interest in visiting the casino, theatre, and chartering a boat.

@@

Although subset 8 was generally less active, almost everyone still engaged in sightseeing. For them, the focus was on sightseeing, relaxing, shopping, and going to the beach, with much less interest in other activities. It can be assumed that they value their time alone. These tourists could potentially be targeted more effectively by offering sightseeing options with minimal interaction, such as using a phone app to provide information about interesting locations, rather than relying on a tour guide.

Finally, since subset 9 is notably inactive, one might infer that they prefer spending much of their time in their accommodation. Consequently, these individuals might be most interested in accommodations that offer well-equipped, comfortable living spaces with amenities that cater to relaxation and leisure.

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{aus_heatmap.png}
    \caption{Heatmap of the intra cluster fraction of the k-means solution with \( k = 6 \) and the reduced feature subset.}
    \label{fig:aus_heatmapg}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{aus_preselection.png}
    \caption{Separation of the data based on the feature "Friends" and the general activity. Top left: 2D tour. Top right: 1D tour. Bottom left: Mosaic plot. Bottom right: Heatmap with the intra cluster fraction.}
    \label{fig:aus_preselection}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{aus_selection.png}
    \caption{Separation of the data based on the feature "Friends" and the general activity after reselection of subsets. 
    Tourists in subsets 7 (pink), 8 (grey) and 9 (gold) all responded thay didn't engage in activities with their friends. Subset 7 was the most active, subset 8 medium active and subset 9 largely inactive. Top left: 2D tour. Top right: 1D tour. Bottom left: Mosaic plot. Bottom right: Heatmap with the intra cluster fraction.}
    \label{fig:aus_selection}
\end{figure}

>>>>

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{silhouette_comparison.png}
    \caption{Comparison of silhouette plots of two k-means cluster solutions of the Austrian vacation activities dataset with k=6. (A) shows the silhouette plot of the k-means solution of the full dataset and (B) the silhouette plot of the k-means solution of the dataset after manual feature selection. We can see that the cluster solution with the reduced dataset achieved better silhouette scores and that clusters 1 and 3 contain observations with negative silhouette scores.}
    \label{fig:silhouette_comparison}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_activ_cluster_highlights.png}
    \caption{Display of a projection of the Austrian vacation activities dataset with the six clusters of the k-means cluster solution highlighted in different colors. The colors indicate which clusters the highlighted observations belong to with cluster 1 being blue (A), cluster 2 being orange (B), cluster 3 being green (C), cluster 4 being red (D), cluster 5 being violett (E) and cluster 6 being brown (F). We can see that the projection separates some clusters well, however also that there is considerable overlap of clusters 4, 5 and 6}
    \label{fig:winter_activ_cluster_highlights}
\end{figure}

This process allows us to visually assess the separation and similarities between the clusters, providing insight into the structure of the dataset. By highlighting each cluster individually, we can evaluate their distinctiveness in different projections. The most influential features shown in Figure \ref{fig:winter_activ_cluster_highlights} are pool/sauna, alpine skiing, museums, and going to the spa. The projection roughly separates clusters 1 (blue), 2 (orange), and 3 (green) from each other and the other three clusters (red, violett and brown), which appear to be quite similar.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_init.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset. Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot. Touirists in both clusters 1 and 3 didn't participate in skiing a lot, but tourists in cluster 3 were much more interested in going to the pool, spa and health facilities compared to cluster 1.}
    \label{fig:winter_cl7_init}
\end{figure}

@@

This setup produces the display shown in Figure \ref{fig:winter_cl7_init}. It can also be seen that both clusters 1 and 3 contain tourists that didn't go alpine skiing and the main difference between them is that tourists in cluster 3 enjoyed going to the spa and health facilities as well as going to the pool, while the ones in cluster 1 didn't. Other than that, the clusters were mostly similar. Now we might be interested in the subset of tourists that enjoy going to museums. Therefore, we can adjust the projection axes so that these axes are elongated and point into different directions. We can see that there is indeed overlap between clusters 1 and 3, as shown in Figure \ref{fig:winter_cl7_pre}. As a next step we can reassign the overlapping section to a new cluster - cluster 7 (pink). By selecting the checkbox for subset 7 and manually selecting the region of overlap, we can form a new cluster, which is visualized in Figure \ref{fig:winter_cl7_post}.

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_pre.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset, with manually adjusted projections. Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot. Changing the projeciton axes of "going to a spa", "museums", "sight seeing" and "using health facilities" reveals the preferences and overlap of clusters 1 and 3.}
    \label{fig:winter_cl7_pre}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winter_cl7_post.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset, after new subsets have been selected. Top left: 2D tour with the linear discriminant analysis projection pursuit index. Top right: heatmap with the intra-cluster fraction. Bottom left: 1D tour with the linear discriminant analysis projection pursuit index. Bottom right: mosaic plot. We can see that now almost all museum goers are in the new manually selected subset 7 and what their preferences are compared to the other clusters.}
    \label{fig:winter_cl7_post}
\end{figure}

@@

At first, hierarchical clustering using Ward's method \citep{murtagh2014ward} and the Jaccard index was applied to the features. The resulting dendrogram is shown in Figure \ref{fig:aus_feature_clustering}. Based on this clustering, \( k = 15 \) clusters were identified, and generally, only one representative feature from each cluster was selected for further analysis. Clusters containing unpopular activities, such as "Adventure", which only had 42 participants, were discarded. The cluster containing popular features like "Beach," "Swimming," "ScenicWalks," "Markets," "Sightseeing," "Friends," "Pubs," "BBQ," "Shopping," "Eating," "EatingHigh," "Movies," and "Relaxing" was treated differently. Multiple features from this cluster were retained to preserve as much information as possible. After feature selection, the observations were reclustered using k-means with \( k = 15 \).

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{aus_feature_clustering.png}
    \caption{Dendogram of the features of the Australian Vacation Activities dataset using Ward's method with the Jaccard index. Features which were clustered together are marked by the colored boxes. We can see, which activties had similar patterns in tourst interest.}
    \label{fig:aus_feature_clustering}
\end{figure}

@@

Since we are particularly interested in tourists who did not spend time with friends, we can extend the projection axis "Friends" outward to separate the data based on that feature. Datapoints in the opposite direction of the "Friends" axes are the solo travellers we are interested in, which can be seen in Figure \ref{fig:aus_preselection} on the left side of the top left plot. Subsequently, we can pull all other projection axes in one direction to separate data points based on their overall activity level. Tourists that fall in the direction of the axes generally engaged in more activities compared to those in the opposite direction of the projection axes. This separation is evident, as observations in subset 6 (brown) are located opposite the axes, and the heatmap shows that they did not engage in many activities. Similarly we can also see that subset 2 (orange), which contains quite active tourists, is shifted towards the direction of the axes.

Using this logic, we can segment the data points on the left into three segments: active tourists (more upward), moderately active tourists (center), and largely inactive tourists (bottom). The result of this segmentation can be seen in Figure \ref{fig:aus_selection}. Analyzing the heatmap, we can identify several interesting patterns.

By comparing subset 3 (green), which contains active tourists who spent time with their friends, with subset 7 (pink), the active tourists who traveled alone, we notice that subset 7 showed less interest in visiting the casino, theatre, and chartering a boat.

@@

Although subset 8 (grey) was generally less active, almost everyone still engaged in sightseeing. For them, the focus was on sightseeing, relaxing, shopping, and going to the beach, with much less interest in other activities. It can be assumed that they value their time alone. These tourists could potentially be targeted more effectively by offering sightseeing options with minimal interaction, such as using a phone app to provide information about interesting locations, rather than relying on a tour guide.

Finally, since subset 9 (gold) is notably inactive, one might infer that they prefer spending much of their time in their accommodation. Consequently, these individuals might be most interested in accommodations that offer well-equipped, comfortable living spaces with amenities that cater to relaxation and leisure.

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{aus_heatmap.png}
    \caption{Heatmap of the intra cluster fraction (indicated by color) of the k-means solution with \( k = 6 \) and the reduced feature subset. We can observe that tourists in subsets 1, 2 and 6 all prefer to travel without their friends.}
    \label{fig:aus_heatmapg}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{aus_preselection.png}
    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Australian vacation activities dataset, with the projection axes of the feature "Friends" pointing into one direction and all the other ones into another one. Top left: 2D tour. Top right: 1D tour. Bottom left: Mosaic plot. Bottom right: Heatmap with the intra cluster fraction. We can see how the feature "Friends" and the general activity level (represented by the other projection axes pointing into one direction) separates the data.}
    \label{fig:aus_preselection}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{aus_selection.png}
    \caption{Interactive tour GUI as seen in \ref{fig:aus_preselection}, but after subselection of subsets 7 (pink), 8 (grey) and 9 (gold). Top left: 2D tour. Top right: 1D tour. Bottom left: Mosaic plot. Bottom right: Heatmap with the intra cluster fraction. We can observe the preferences of three new subsegments, which can be interpreted as very active (pink), moderately active (grey) and inactive (gold) tourists tavelling without their friends.}
    \label{fig:aus_selection}
\end{figure}
